Can Symmetric Cryptography Be Liberated from the Von Neumann Style?

### 2024-12-11 paper.tex
Tone down banter
Justify order of bitsliced state functor
Reviewer A:
  p10, l392: je pense qu'un programmeur typique aura l'intuition que
  la définition de `reg32` va être `Double.T (Rows.T (Cols.T A))`,
  mais ce n'est pas le choix que vous faites. Je réalise que c'est
  l'essence du bitslicing, mais comme vous le remarquez dans votre
  introduction, ce n'est pas une technique particulièrement connue
  hors du milieu de la cryptographie. Je pense qu'il faut expliquer
  pourquoi cette représentation est le bon choix.

The role of the diffusion layer is to divert individual bits across the
whole structure. Inevitably, this calls for a notion of indexing.
Naperian functors~\citep{hancock:napierian} identify a type \coqe!Ix! of
indices as the logarithm of an exponential type:
%
\coqfrom{Algebra}{Naperian}
%
through the fact that \coqe!lookup! and \coqe!init! form a bijection.
This is also known as a representable functor in categorical
circles~\citep{maclane:working-categorist}.

@@

Storing only 16 bits in a 32 bits word (on \textsc{Arm} Cortex M)
remains sub-optimal. We can further increase register usage and
instantly double throughput by processing two blocks at the same
time. To this end, we introduce a new Naperian functor,
\coqe!Double.T!, indexed by the two elements set \coqe!Fst!  and
\coqe!Snd!. This leads to
%
\coqfrom{Skinny/Bitslice}{state}
%
which denote our intent to manipulate objects of type \coqe!reg32! as
if they were stored in a single machine word. In the present work,
this observation will remain at the state of wishful thinking. One can
think of the code in this section (as well as the fixsliced code, in
the next section) as playing, themselves, the role of specifications
to lower-level refinements. Our previous work on Usuba~\citep{mercadier:PhD} suggests that
we will be able to deliver on this promise in the future. Further,
remark that the \coqe!state! now corresponds to two intermingled
blocks but as we shall process them in lock-step throughout the
cipher, this does not make much difference.

>>>>

The role of the diffusion layer is to divert individual bits across the
whole structure. Inevitably, this calls for a notion of indexing.
Naperian functors~\citep{hancock:napierian} identify a type \coqe!Ix! of
indices as the logarithm of an exponential type:
%
\coqfrom{Algebra}{Naperian}
%
through the fact that \coqe!lookup! and \coqe!init! form a bijection.
This is also known as a representable functor in the categorical
literature~\citep{maclane:working-categorist}.

@@

Storing only 16 bits in a 32 bits word (on \textsc{Arm} Cortex M)
remains sub-optimal. We can further increase register usage and
instantly double throughput by processing two blocks at the same
time. To this end, we introduce a new Naperian functor,
\coqe!Double.T!, indexed by the two elements set \coqe!Fst!  and
\coqe!Snd!. Since \coqe!Double.T bool! is a Boolean algebra, we immediately get that \coqe!subcells_! and \coqe!add_shift_rows_! can process two blocks at the same time. This leads to
%
\coqfrom{Skinny/Bitslice}{state}
%
which denote our intent to manipulate objects of type \coqe!reg32! as
if they were stored in a single machine word.  In the present work,
this observation will remain at the state of wishful thinking. One can
think of the code in this section (as well as the fixsliced code, in
the next section) as playing, themselves, the role of specifications
to lower-level refinements. Our previous work on Usuba~\citep{mercadier:PhD} suggests that
we will be able to deliver on this promise in the future. Further,
remark that the \coqe!state! now corresponds to two intermingled
blocks but as we shall process them in lock-step throughout the
cipher, this does not make much difference.
